---
title: 'Codebook: Study on value creation model of science-society interactions'
author: "Irene Ramos-Vielba, Nicolas Robinson-Garcia and Richard Woolley"
date: "March 18, 2022"
output: html_document
toc: TRUE
toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Required libraries
library(archetypes)
library(ggplot2)
library(tidyverse)
library(corrplot)
library(dplyr)
library(Hmisc)
library(GGally)
library(archetypes)

```

# Brief description

This is document reproduces the figures produced in the study 'A value creation model for science-society interconnections: Archetypal analysis combining publications, survey and altmetric data' authored by I. Ramos-Vielba, N. Robinson-Garcia /& R. Woolley. Data is available at [doi:10.5281/zenodo.6393226](https://doi.org/10.5281/zenodo.6393226). Code scripts are available in a (GitHub repository)[https://github.com/elrobin/value-components].

# Data import and processing

Data is imported from Zenodo. It includes 9,190 observations and 12 variables.

```{r}
df <- read.csv(
  "https://zenodo.org/record/6393226/files/dataset_archetypes_v3.csv?download=1", sep = ";"
)
str(df)
```

We create non-size dependent variables based on publications by converting these into proportions and we then rename them to match the naming of the value creation model.

```{r}
df$pp_news <- df$p_news / df$PUBS
df$pp_policy <- df$p_policy / df$PUBS
df$pp_oa <- df$OA / df$PUBS
df$pp_twitter <- df$p_twitter / df$PUBS

df <- df %>% # Select and rename variables
  select(
    ï..id,
    translation,
    dissemination,
    engagement,
    production,
    pp_news,
    pp_oa,
    pp_policy,
    pp_twitter,
    transmission,
    field
  )

names(df) <- c(
  "ID_FINAL",
  "Commercialisation",
  "Dissemination",
  "Engagement",
  "Joint research",
  "Media promotion",
  "Openness",
  "Public Policy",
  "Social visibility",
  "Transmission",
  "field"
)
```

# Figures

## Descriptives
The following table corresponds to Table 5 in the paper, and includes basic descriptives per field:

- BIOMED. Biomedicine
- SSH. Social Sciences and Humanities
- STEM. Science, Technology, Engineering and Mathematics.

```{r}
df.vars <- df %>%
  tidyr::pivot_longer(!field & !ID_FINAL, 
                      names_to = "variable",
                      values_to = "value")

df.vars %>%
  group_by(field, variable) %>%
summarise(n = n(), mean = mean(value), sd = sd(value))
```

## Correlation matrix

Next we create a correlation matrix including also the distribution of each variable. This figure corresponds to Figure 2.

```{r}
  ggpairs(df, columns=2:10, aes(colour = field)) + 
    theme_bw() + scale_color_brewer(palette = "Dark2")
```

## Archetypal analyses

Following we include the archetypal analyses constructed for each field and overall.

### All cases

The archetypal analysis for all cases is displayed in the manuscript in Figure 3.

#### Identification of archetypes

We first identify the ideal number of archetypes needed. We search for the best model looking at up to 5 archetypes and 5 repetitions.

```{r}
all <- df %>%
  select(!field & !ID_FINAL)

set.seed(1986)
arc.k <- stepArchetypes(data = all, 
                        k = 1:5, 
                        verbose =F, 
                        nrep = 5)
```

The optimal number of archetypes is selected based on the Residual Sum of Squares (RSS) for which we apply an elbow rule.

```{r}
screeplot(arc.k) # 3 archetypes
```

#### Characterization of archetypes

These archetypes have the following parameters

```{r}
arc.model <- bestModel(arc.k[[3]])
t(parameters(arc.model))
```

Which can be better interpreted when plotted in percentiles:

```{r}

barplot(arc.model, all, percentiles = TRUE)
```

In the manuscript these percentiles are shown using a radar chart for which we first have to recalculate those percentiles as computed in the `archetypes` package.

```{r}
.perc <- function(x, data, digits = 2) { # Calculates percentiles
  Fn <- ecdf(data)
  round(Fn(x) * 1, digits = digits) 
}

param <- parameters(arc.model)


atypes <- param # extract parameters
rownames(atypes) <- sprintf('Archetype %s',
                            seq(length = nrow(atypes)))

atypes <- sapply(seq(length = ncol(all)), # Identify percentile for param
                 function(i)              # based on a data.frame
                   .perc(atypes[, i], all[, i]))

colnames(atypes) <- colnames(all)
atypes <- as.data.frame(atypes)

atypes$archetype <- c(1,2,3)
```

We can then produce the radar charts after converting from wide to long format

```{r}
params <-
  tidyr::pivot_longer(atypes,
                      !c("archetype"),
                      names_to = "variable",
                      values_to = "value")
a1 <- 
  ggplot(subset(params, archetype == 1)) +
  geom_col(aes(x = variable, y = value), fill = "#F25200") +
  ylim(0, 1) +
  labs(title = "Archetype 1", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()

a2 <-
  ggplot(subset(params, archetype == 2)) +
  geom_col(aes(x = variable, y = value), fill = "#F25200") +
  ylim(0, 1) +
  labs(title = "Archetype 2", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()
  
a3 <-
  ggplot(subset(params, archetype == 3)) +
  geom_col(aes(x = variable, y = value), fill = "#F25200") +
  ylim(0, 1) +
  labs(title = "Archetype 3", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()

par(mfrow = c(3, 1))
a1
a2
a3
```

Co-occurrence term maps overlaying alpha scores for researchers' scientific output is available online here: (https://sl.ugr.es/allarchetypes)[https://sl.ugr.es/allarchetypes]. Below find the code to extract the alpha score of each observation for the three archetypes:

```{r}
# alphas <- arc.model[["alphas"]]
# 
# alphas <- cbind.data.frame(df$ID_FINAL, alphas)
# write.csv(alphas, file = "alpha-all-respondents.csv", quote = F, 
#            row.names = F,)
```

### STEM

The archetypal analysis for STEM cases is displayed in the manuscript in Figure 4.

#### Identification of archetypes

We first identify the ideal number of archetypes needed. We search for the best model looking at up to 5 archetypes and 5 repetitions.

```{r}
stem <- df %>%
  filter(field=="STEM") %>%
  select(!field & !ID_FINAL)

set.seed(1986)
arc.k <- stepArchetypes(data = stem, 
                        k = 1:5, 
                        verbose =F, 
                        nrep = 5)
```

The optimal number of archetypes is selected based on the Residual Sum of Squares (RSS) for which we apply an elbow rule.

```{r}
screeplot(arc.k) # 3 archetypes
```

#### Characterization of archetypes

These archetypes have the following parameters

```{r}
arc.model <- bestModel(arc.k[[3]])
t(parameters(arc.model))
```

Which can be better interpreted when plotted in percentiles:

```{r}
barplot(arc.model, stem, percentiles = TRUE)
```

In the manuscript these percentiles are shown using a radar chart for which we first have to recalculate those percentiles as computed in the `archetypes` package.

```{r}
stemtypes <- parameters(arc.model) # extract parameters
rownames(stemtypes) <- sprintf('Archetype %s',
                               seq(length = nrow(stemtypes)))

stemtypes <-
  sapply(seq(length = ncol(stem)), # Identify percentile for param
         function(i)
           # based on a data.frame
           .perc(stemtypes[, i], stem[, i]))

colnames(stemtypes) <- colnames(stem)
stemtypes <- as.data.frame(stemtypes)

stemtypes$archetype <- c(1, 2, 3)
```

We can then produce the radar charts after converting from wide to long format

```{r}
params <-
  tidyr::pivot_longer(stemtypes,
                      !c("archetype"),
                      names_to = "variable",
                      values_to = "value")
a1 <- 
  ggplot(subset(params, archetype == 1)) +
  geom_col(aes(x = variable, y = value), fill = "#00BAB4") +
  ylim(0, 1) +
  labs(title = "Archetype STEM1", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()

a2 <-
  ggplot(subset(params, archetype == 2)) +
  geom_col(aes(x = variable, y = value), fill = "#00BAB4") +
  ylim(0, 1) +
  labs(title = "Archetype STEM2", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()
  
a3 <-
  ggplot(subset(params, archetype == 3)) +
  geom_col(aes(x = variable, y = value), fill = "#00BAB4") +
  ylim(0, 1) +
  labs(title = "Archetype STEM3", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()

par(mfrow = c(3, 1))
a1
a2
a3
```

Co-occurrence term maps overlaying alpha scores for researchers' scientific output is available online here: (https://sl.ugr.es/stemarchetypes)[https://sl.ugr.es/stemarchetypes].
Below find the code to extract the alpha score of each observation for the three archetypes:

```{r}
alphas <- arc.model[["alphas"]]

df.stem <- subset (df,
                   field=="STEM",
                   select =
                     c("ID_FINAL"))
alphas <- cbind.data.frame(df.stem, alphas)

write.csv(alphas, file = "alpha-stem-respondents.csv", quote = F, 
          row.names = F)
```

### SSH

The archetypal analysis for all cases is displayed in the manuscript in Figure 5.

#### Identification of archetypes

We first identify the ideal number of archetypes needed. We search for the best model looking at up to 5 archetypes and 5 repetitions.

```{r}
ssh <- df %>%
  filter(field=="SSH") %>%
  select(!field & !ID_FINAL)

set.seed(1986)
arc.k <- stepArchetypes(data = ssh, 
                        k = 1:5, 
                        verbose =F, 
                        nrep = 5)
```

The optimal number of archetypes is selected based on the Residual Sum of Squares (RSS) for which we apply an elbow rule.

```{r}
screeplot(arc.k) # 3 archetypes
```

#### Characterization of archetypes

These archetypes have the following parameters

```{r}
arc.model <- bestModel(arc.k[[3]])
t(parameters(arc.model))
```

Which can be better interpreted when plotted in percentiles:

```{r}
barplot(arc.model, stem, percentiles = TRUE)
```

In the manuscript these percentiles are shown using a radar chart for which we first have to recalculate those percentiles as computed in the `archetypes` package.

```{r}
sshtypes <- parameters(arc.model) # extract parameters
rownames(sshtypes) <- sprintf('Archetype %s',
                               seq(length = nrow(sshtypes)))

sshtypes <-
  sapply(seq(length = ncol(ssh)), # Identify percentile for param
         function(i)
           # based on a data.frame
           .perc(sshtypes[, i], ssh[, i]))

colnames(sshtypes) <- colnames(ssh)
sshtypes <- as.data.frame(sshtypes)

sshtypes$archetype <- c(1, 2, 3)
```

We can then produce the radar charts after converting from wide to long format

```{r}
params <-
  tidyr::pivot_longer(sshtypes,
                      !c("archetype"),
                      names_to = "variable",
                      values_to = "value")
a1 <- 
  ggplot(subset(params, archetype == 1)) +
  geom_col(aes(x = variable, y = value), fill = "#304160") +
  ylim(0, 1) +
  labs(title = "Archetype SSH1", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()

a2 <-
  ggplot(subset(params, archetype == 2)) +
  geom_col(aes(x = variable, y = value), fill = "#304160") +
  ylim(0, 1) +
  labs(title = "Archetype SSH2", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()

a3 <-
  ggplot(subset(params, archetype == 3)) +
  geom_col(aes(x = variable, y = value), fill = "#304160") +
  ylim(0, 1) +
  labs(title = "Archetype SSH3", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()

par(mfrow = c(3, 1))
a1
a2
a3

```

Co-occurrence term maps overlaying alpha scores for researchers' scientific output is available online here: (https://sl.ugr.es/ssharchetypes)[https://sl.ugr.es/ssharchetypes].

Below find the code to extract the alpha score of each observation for the three archetypes:

```{r}
alphas <- arc.model[["alphas"]]

df.ssh <- subset (df,
                   field=="SSH",
                   select =
                     c("ID_FINAL"))
alphas <- cbind.data.frame(df.ssh, alphas)

write.csv(alphas, file = "alpha-ssh-respondents.csv", quote = F, 
          row.names = F)
```

### BIOMED

The archetypal analysis for BIOMED cases is displayed in the manuscript in Figure 6.

#### Identification of archetypes

We first identify the ideal number of archetypes needed. We search for the best model looking at up to 5 archetypes and 5 repetitions.

```{r}
biomed <- df %>%
  filter(field=="BIOMED") %>%
  select(!field & !ID_FINAL)

set.seed(1986)
arc.k <- stepArchetypes(data = biomed, 
                        k = 1:5, 
                        verbose =F, 
                        nrep = 5)
```

The optimal number of archetypes is selected based on the Residual Sum of Squares (RSS) for which we apply an elbow rule.

```{r}
screeplot(arc.k) # 3 archetypes
```

#### Characterization of archetypes

These archetypes have the following parameters

```{r}
arc.model <- bestModel(arc.k[[3]])
t(parameters(arc.model))
```

Which can be better interpreted when plotted in percentiles:

```{r}
barplot(arc.model, biomed, percentiles = TRUE)
```

In the manuscript these percentiles are shown using a radar chart for which we first have to recalculate those percentiles as computed in the `archetypes` package.

```{r}
biomedtypes <- parameters(arc.model) # extract parameters
rownames(biomedtypes) <- sprintf('Archetype %s',
                               seq(length = nrow(biomedtypes)))

biomedtypes <-
  sapply(seq(length = ncol(biomed)), # Identify percentile for param
         function(i)
           # based on a data.frame
           .perc(biomedtypes[, i], biomed[, i]))

colnames(biomedtypes) <- colnames(biomed)
biomedtypes <- as.data.frame(biomedtypes)

biomedtypes$archetype <- c(1, 2, 3)
```

We can then produce the radar charts after converting from wide to long format

```{r}
params <-
  tidyr::pivot_longer(biomedtypes,
                      !c("archetype"),
                      names_to = "variable",
                      values_to = "value")
a1 <- 
  ggplot(subset(params, archetype == 1)) +
  geom_col(aes(x = variable, y = value), fill = "#ADA16B") +
  ylim(0, 1) +
  labs(title = "Archetype BIOMED1", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()

a2 <-
  ggplot(subset(params, archetype == 2)) +
  geom_col(aes(x = variable, y = value), fill = "#ADA16B") +
  ylim(0, 1) +
  labs(title = "Archetype BIOMED2", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()
  
a3 <-
  ggplot(subset(params, archetype == 3)) +
  geom_col(aes(x = variable, y = value), fill = "#ADA16B") +
  ylim(0, 1) +
  labs(title = "Archetype BIOMED3", 
       x = element_blank(), 
       y = element_blank()) +
  coord_polar() +
  theme_minimal()

par(mfrow = c(3, 1))
a1
a2
a3
```

Co-occurrence term maps overlaying alpha scores for researchers' scientific output is available online here: (https://sl.ugr.es/biomedarchetypes)[https://sl.ugr.es/biomedarchetypes].
Below find the code to extract the alpha score of each observation for the three archetypes:

```{r}
alphas <- arc.model[["alphas"]]

df.biomed <- subset (df,
                   field=="BIOMED",
                   select =
                     c("ID_FINAL"))
alphas <- cbind.data.frame(df.biomed, alphas)

write.csv(alphas, file = "alpha-biomed-respondents.csv", quote = F, 
          row.names = F)
```
